Formal Modeling and Verification of a Rate-Monotonic Scheduling Implementation with Real-Time Maude

### 2016-10-26 RMS/final-version/main.tex
modified fig rms

A simple example showing this algorithm is depicted in
Figure~\ref{sf:rmsalg}. However in this paper, we consider an
implementation instead of the RMS algorithm itself, thus the model
would be more complicated than this standard, ideal
setting. Assumption~(A1) will be modified because of the interrupt
mask mechanism, while (A4) is relaxed to obtain a more realistic
analysis model.

@@

These assumptions make our model different from the standard one. For
instance, with assumption~(A1'), an interrupt request that occurs
during switching will be pending, such that jobs of task $\tau_i$
cannot initiate at $kT_i$. They should wait until switching finishes
and the interrupt mask bit is cleared, which is different
from~(A1). Note that (A3) says that jobs are ready to run at their
initiation times, however, no one can start running at exactly its
initiation time, because scheduling takes time. Under these
assumptions, an example showing the execution of tasks scheduled by
the implementation is depicted in Figure~\ref{sf:rmsimp}. Note that
the second job instance of $\tau_1$ is initiated at time $11$ instead
of $10$, because switching is performed and the interrupt mask bit is
set at time $10$, exemplifying differences between (A1') and (A1).

\begin{figure}[!t]
\centering
\subfloat[RMS Algorithm]{
  \begin{tikzpicture}[font=\small,scale=1.25,
    % styles
    system task/.style={fill=gray!70,draw=gray},
    periodic task/.style={fill=gray!70!black,draw=gray!50!black},
    task line/.style={gray!30,very thin},
    time line/.style={gray!50,very thin,dashed},
    short line/.style={gray!50,very thin,dashed},
    init/.style={{Triangle[fill=white,scale=1.1]}-},
    complete/.style={{Triangle[scale=1.1]}-}]

  % legend
  \begin{scope}[xshift=-2cm]  
    \draw [fill=gray!25,draw=gray!50,dashed] (1.5,2.3) rectangle +(4,-1);
    \draw (1.5,2.3) node [below right] {LEGEND};
    \draw [init] (2,1.85) -- +(0,-0.01);
    \draw (2.1,2) node [below right] {: Initiation of Tasks};
    \draw [complete] (2,1.42) -- +(0,0.1);
    \draw (2.1,1.7) node [below right] {: Completion of Tasks};
  \end{scope}

    % grid
    \foreach \x in {0.3,0.7}
      \draw[task line] (0,\x) -- (5.2,\x);

    % axes
    \draw[->,thick] (0,0) node [below] {$0$} -- (5.3,0) node [above] {time};
    \draw[thin] (0,0) -- (0,1);

    % time indicator
    \draw [time line] (0.75,0.7) -- (0.75,0);
    \draw (0.75,0) node [below] {$3$} -- +(0,2pt);
    \draw [time line] (1.25,0.3) -- (1.25,0);
    \draw (1.25,0) node [below] {$5$} -- +(0,2pt);
    \draw [time line] (2.5,0.7) -- (2.5,0);
    \draw (2.5,0) node [below] {$10$} -- +(0,2pt);
    \draw [time line] (3.25,0.7) -- (3.25,0);
    \draw (3.25,0) node [below] {$13$} -- +(0,2pt);

    \draw [time line] (5,0.7) -- (5,0);
    \draw (5,0) node [below] {$20$} -- +(0,2pt);

    % tasks
    \draw [periodic task] (0,0.7) rectangle +(0.75,0.14);
    \draw [periodic task] (0.75,0.3) rectangle +(0.5,0.14);
    \draw [periodic task] (2.5,0.7) rectangle +(0.75,0.14);

    % text
    \draw (0,0.75) node [left] {$\tau_1$};
    \draw (0,0.35) node [left] {$\tau_2$};
    \path (0,0.75) node [left,text opacity=0] {scheduling};
    \path (0,0.35) node [left,text opacity=0] {switching};

    % initiation of tasks
    \foreach \x in {0,2.5,5}
      \draw [init] (\x,0.7) -- +(0,-0.01);
    \foreach \x in {0,5}
      \draw [init] (\x,0.3) -- +(0,-0.01);

    % completion of tasks
    \draw [complete] (0.75,0.84) -- +(0,0.01);
    \draw [complete] (3.25,0.84) -- +(0,0.01);
    \draw [complete] (1.25,0.44) -- +(0,0.01);
  \end{tikzpicture}
\label{sf:rmsalg}}\\
\subfloat[RMS Implementation]{
  \begin{tikzpicture}[font=\small,scale=1.25,
    % styles
    system task/.style={fill=gray!70,draw=gray},
    periodic task/.style={fill=gray!70!black,draw=gray!50!black},
    task line/.style={gray!30,very thin},
    time line/.style={gray!50,very thin,dashed},
    short line/.style={gray!50,very thin,dashed},
    init/.style={{Triangle[fill=white,scale=1.1]}-},
    complete/.style={{Triangle[scale=1.1]}-}]

    % grid
    \foreach \x in {0.3,0.7,1.1,1.5}
      \draw[task line] (0,\x) -- (5.2,\x);

    % axes
    \draw[->,thick] (0,0) node [below] {$0$} -- (5.3,0) node [above] {time};
    \draw[thin] (0,0) -- (0,1.8);

    % time indicator
    \draw [time line] (0.5,1.5) -- (0.5,0);
    \draw (0.5,0) node [below] {$2$} -- +(0,2pt);
    \draw [time line] (1.25,1.5) -- (1.25,0);
    \draw (1.25,0) node [below] {$5$} -- +(0,2pt);
    \draw [time line] (1.75,1.1) -- (1.75,0);
    \draw (1.75,0) node [below] {$7$} -- +(0,2pt);
    \draw [time line] (2.25,1.1) -- (2.25,0);
    \draw (2.25,0) node [below] {$9$} -- +(0,2pt);
    \draw [time line] (2.75,1.5) -- (2.75,0);
    \draw (2.75,0) node [below] {$11$} -- +(0,2pt);
    \draw [time line] (3.25,1.5) -- (3.25,0);
    \draw (3.25,0) node [below] {$13$} -- +(0,2pt);
    \draw [time line] (4,1.5) -- (4,0);
    \draw (4,0) node [below] {$16$} -- +(0,2pt);
    \draw [time line] (4.5,0.3) -- (4.5,0);
    \draw (4.5,0) -- +(0,2pt);
    \draw (4.5,0) node [below] {$18$};
    \draw [time line] (5,1.5) -- (5,0);
    \draw (5,0) node [below] {$20$} -- +(0,2pt);

    % tasks
    \draw [system task] (0,0.7) rectangle +(0.5,0.14);
    \draw [periodic task] (0.5,1.5) rectangle +(0.75,0.14);
    \draw [system task] (1.25,0.3) rectangle +(0.5,0.14);
    \draw [periodic task] (1.75,1.1) rectangle +(0.5,0.14);
    \draw [system task] (2.25,0.3) rectangle +(0.5,0.14);
    \draw [system task] (2.75,0.7) rectangle +(0.5,0.14);
    \draw [periodic task] (3.25,1.5) rectangle +(0.75,0.14);
    \draw [system task] (4,0.3) rectangle +(0.5,0.14);

    % text
    \draw (0,1.55) node [left] {$\tau_1$};
    \draw (0,1.15) node [left] {$\tau_2$};
    \draw (0,0.75) node [left] {scheduling};
    \draw (0,0.35) node [left] {switching};

    % initiation of tasks
    \foreach \x in {0,2.75,5}
      \draw [init] (\x,1.5) -- +(0,-0.01);
    \foreach \x in {0,5}
      \draw [init] (\x,1.1) -- +(0,-0.01);

    % completion of tasks
    \draw [complete] (1.25,1.64) -- +(0,0.01);
    \draw [complete] (2.25,1.24) -- +(0,0.01);
    \draw [complete] (4,1.64) -- +(0,0.01);
  \end{tikzpicture}
\label{sf:rmsimp}}
\caption{A set of $2$ periodic tasks scheduled under RMS algorithm and
  the implementation, respectively. $T=T_1=10$, $T_2=20$, $C_1=3$,
  $C_2=2$. In Figure~\ref{sf:rmsimp}, we assume both scheduling time
  and switching time are $2$.}
\label{f:example}
\end{figure}

@@

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{J_Liu.jpg}}]%
{Jiaxiang Liu} received his B.S. degree from the School of Software,
Tsinghua University, China, in 2010. He is now a Ph.D. candidate in
the Department of Computer Science and Technology, Tsinghua
University. His research interests include formal methods, rewrite
systems and embedded systems.
\end{IEEEbiography}

>>>>

A simple example showing this algorithm is depicted in
Fig.~\ref{f:example}(a). However in this paper, we consider an
implementation instead of the RMS algorithm itself, thus the model
would be more complicated than this standard, ideal
setting. Assumption~(A1) will be modified because of the interrupt
mask mechanism, while (A4) is relaxed to obtain a more realistic
analysis model.

@@

These assumptions make our model different from the standard one. For
instance, with assumption~(A1'), an interrupt request that occurs
during switching will be pending, such that jobs of task $\tau_i$
cannot initiate at $kT_i$. They should wait until switching finishes
and the interrupt mask bit is cleared, which is different
from~(A1). Note that (A3) says that jobs are ready to run at their
initiation times, however, no one can start running at exactly its
initiation time, because scheduling takes time. Under these
assumptions, an example showing the execution of tasks scheduled by
the implementation is depicted in Fig.~\ref{f:example}(b). Note that
the second job instance of $\tau_1$ is initiated at time $11$ instead
of $10$, because switching is performed and the interrupt mask bit is
set at time $10$, exemplifying differences between (A1') and (A1).

\begin{figure}[!t]
\centering
\includegraphics{rms.pdf}
\caption{A set of $2$ periodic tasks scheduled under RMS algorithm and
  the implementation, respectively. $T=T_1=10$, $T_2=20$, $C_1=3$,
  $C_2=2$. In~(b), we assume both scheduling time and switching time
  are $2$.}
\label{f:example}
\end{figure}

@@

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{J_Liu.jpg}}]%
{Jiaxiang Liu} received his B.S. degree from the School of Software,
Tsinghua University, China, in 2010. He is now a co-supervised
Ph.D. candidate between the Department of Computer Science and
Technology, Tsinghua University, China and LIX, \'Ecole Polytechnique,
France. His research interests include formal methods, rewrite systems
and embedded systems.
\end{IEEEbiography}
